Characterizing Water Composition with an Autonomous Robotic Team Employing Comprehensive In Situ Sensing, Hyperspectral Imaging, Machine Learning, and Conformal Prediction

### 2024-03-11 paper/figures/results/maps/supplementary-material/spectral-indices.qmd
final updates

\begin{align}\label{eq:ref-bands}
\begin{split}
    R_b &= R(440 \text{ nm}) \\
    R_g &= R(550 \text{ nm}) \\
    R_b &= R(650 \text{ nm}) \\
    R_{nir} &= R(860 \text{ nm}).
\end{split}
\end{align}
We also define $R_{swir} = R(1009 \text{ nm})$ as the *almost*-SWIR band representing the infrared edge of the hyperspectral imager.

@@

| Normalized Difference Water Index 2 | NDWI2 | $\dfrac{R_{nir} - R_{swir}}{R_{nir} + R_{swir}}$ |
| Modified Normalized Difference Water Index | MNDWI | $\dfrac{R_g - R_{swir}}{R_g + R_{swir}}$ |

@@

| Moisture Stress Index | MSI | $\dfrac{R_{swir}}{R_{nir}}$ |

@@

| Normalzied Difference Infrared Index | NDII | $\dfrac{R_{nir} - R_{swir}}{R_{nir} + R_{swir}}$ |

>>>>

\begin{align}\label{eq:ref-bands}
\begin{split}
    R_b &= R(440 \text{ nm}) \\
    R_g &= R(550 \text{ nm}) \\
    R_b &= R(650 \text{ nm}) \\
    R_{nir} &= R(860 \text{ nm}).
\end{split}
\end{align}
We also define $R_{far} = R(1009 \text{ nm})$ as the band representing the farthest infrared bin for our hyperspectral imager which we use instead of the usual SWIR band.

@@

| Normalized Difference Water Index 2 | NDWI2 | $\dfrac{R_{nir} - R_{far}}{R_{nir} + R_{far}}$ |
| Modified Normalized Difference Water Index | MNDWI | $\dfrac{R_g - R_{far}}{R_g + R_{far}}$ |

@@

| Moisture Stress Index | MSI | $\dfrac{R_{far}}{R_{nir}}$ |

@@

| Normalzied Difference Infrared Index | NDII | $\dfrac{R_{nir} - R_{far}}{R_{nir} + R_{far}}$ |

### 2024-03-11 paper/remotesensing-2870812-english.tex
final updates

%=================================================================
% Full title of the paper (Capitalized)
 \Title{Characterizing Water Composition with an Autonomous Robotic Team Employing Comprehensive In Situ Sensing, Hyperspectral Imaging, Machine Learning, and Conformal Prediction}%AE title change
%\Title{\textcolor{red}{Towards Real-Time Water Quality Characterization Using an Autonomous Robotic Team Employing Comprehensive In-Situ Sensing, Hyperspectral Imaging, and Conformal Prediction}}

@@

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
Hanson Center for Space Sciences, University of Texas at Dallas, Richardson, TX 75080, USA;  john.waczak@utdallas.edu (J.W.); adam.aker@utdallas.edu (A.A.); lhw150030@utdallas.edu (L.O.H.W.); shawhintalebi@gmail.com (S.T.); ashen.fernando@utdallas.edu  (B.F.); pxh180012@utdallas.edu (P.M.H.D.); mazhar.iqbal@utdallas.edu (M.I.); matthew.lary@utdallas.edu (M.L.); captdaveschaefer@gmail.com (D.S.)}

@@

\begin{figure}[H]
\vspace{-0.15in}
\includegraphics[width=0.9\columnwidth]{figures/results/fits/physical-fitres.png}
\vspace{-0.1in}
\caption{\hl{Scatter} %MDPI: 1. Please change hyphen into endash for range. 2. Please use commas to separate thousands for numbers with five or more digits (not four digits) in the picture, e.g., "10000" should be "10,000". 
 diagrams (\textbf{left}) and quantile--quantile plots (\textbf{right}) for the hyperparameter-optimized RFR models for the physical variables measured by the USV.\label{fig:physical-fit}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/fits/physical-ranking.pdf}
\vspace{-0.1in}
\caption{Ranked
 \hl{permutation} %MDPI: Please change the hyphen (-) into a minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 importance for each feature in the physical variable models. Permutation importance measured the decrease in the model's $R^2$ value after replacing each feature in the prediction set with a random permutation of its values.\label{fig:physical-fi}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=0.9\columnwidth]{figures/results/maps/physical.pdf}
\vspace{-0.05in}
\caption{\hl{Maps} %MDPI: Please change the hyphen (-) into a minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 generated by applying each of the physical variable models to the hyperspectral data cubes collected on 23 November. Overlaid over the predictions are color-filled squares showing the associated in situ reference data for the same collection period. The size of the squares has been exaggerated for visualization. We note that there is good agreement between the model predictions and the reference data. \label{fig:map-physical}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/fits/ions-fitres.png}
\vspace{-0.1in}
\caption{\hl{Scatter} %MDPI: Please use commas to separate thousands for numbers with five or more digits (not four digits) in the picture, e.g., "10000" should be "10,000".
 diagrams (\textbf{left}) and quantile--quantile plots (\textbf{right}) for the hyperparameter-optimized RFR models for the ion measurements made by the USV.\label{fig:ions-fit}}
\end{figure}

\vspace{-9pt}
\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\vspace{-0.1in}
\includegraphics[width=18 cm]{figures/results/fits/ions-ranking.pdf}
\vspace{-0.1in}
\end{adjustwidth}
\caption{\hl{Ranked} %MDPI: Please change the hyphen (-) into a minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 permutation importance for the top 25 features of the ion models. The permutation importance measures the decrease in the model's $R^2$ value when each feature is replaced by a random permutation of its values.\label{fig:ions-fi}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/maps/ions.pdf}
\vspace{-0.1in}
\caption{\hl{Maps} %MDPI: Please change the hyphen (-) into a minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 generated by applying the trained ion models to the data cubes collected on 23 November. Overlaid on the maps are the in situ reference measurements for the same collection period. The size of the squares has been exaggerated for the visualization. We note that there is good agreement between the generated map and the reference data. \label{fig:map-ions}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/fits/biochemical-fitres.png}
\vspace{-0.1in}
\caption{\hl{Scatter} %MDPI: Please use commas to separate thousands for numbers with five or more digits (not four digits) in the picture, e.g., "10000" should be "10,000".
 plots (\textbf{left}) and quantile--quantile plots (\textbf{right}) for the final hyperparameter-optimized models for the biochemical targets blue--green algae (phycoerythrin), CDOM, chlorophyll-a, and blue--green algae (phycocyanin). \label{fig:biochem-fit}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=0.9\columnwidth]{figures/results/fits/biochemical-ranking.pdf}
\vspace{-0.1in}
\caption{\hl{Ranked} %MDPI: Please change the hyphen (-) into a minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 permutation importance for each feature in the trained biochemical models. The permutation importance measures the decrease in the model's $R^2$ value after replacing each feature with a random permutation of its values.\label{fig:biochem-fi}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=0.92\columnwidth]{figures/results/maps/biochemical.pdf}
\vspace{-0.1in}
\caption{\hl{Maps} %MDPI: Please change the hyphen (-) into a minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 \hl{generated} %MDPI: There are overlapped content (``longitude} in top and left subfigure). please revise.
 by applying the trained biochemical models to the data cubes collected on 23 November. Overlaid are the in situ reference data for the same collection period. The size of the squares has been exaggerated for the visualization. We note there is good agreement between the predicted map and the reference data. \label{fig:map-biochem}}
\end{figure}

@@

The final two models to consider are for the measured chemical concentrations of crude oil (CO) and optical brighteners (OB). The crude oil measurement includes natural unprocessed petroleum, whereas optical brighteners consist of whitening agents that are often added to products such as soaps, detergents, and cleaning agents. Both the crude oil and optical brightener measurements show multi-modal spatial distributions across each collection period. Scatter diagrams and quantile--quantile plots for the fitted models are shown in Figure~\ref{fig:chemicals-fit}. Both models achieve good performance, with $R^2$ values of 0.957 and 0.941 for CO and OB, respectively, %Please ensure meaning has been retained. 
on the holdout test set. The performance of the CO model degrades for concentrations below 24 ppb, for which there are few records. Similarly, the OB model shows worse performance for concentrations below roughly 3.5 ppb.

@@

\vspace{-0.21in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/fits/chemical-fitres.png}
\vspace{-0.11in}
\caption{\hl{Scatter} %MDPI: Please use commas to separate thousands for numbers with five or more digits (not four digits) in the picture, e.g., "10000" should be "10,000".
 diagrams (\textbf{left}) and quantile--quantile plots (\textbf{right}) for the hyperparameter-optimized RFR models for the chemical variables measured by the USV.\label{fig:chemicals-fit}}
\end{figure}

\vspace{-9pt}
\begin{figure}[H]
\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/fits/chemical-ranking.pdf}
\vspace{-0.15in}
\caption{\hl{Ranked} %MDPI: Please change the hyphen (-) into a minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 permutation importance for the top 25 features of the chemical models. The permutation importance measures the decrease in the model's $R^2$ value after replacing each feature in the prediction set with a random permutation of its values.\label{fig:chemicals-fi}}
\end{figure}

@@

\vspace{-0.2in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/maps/chemical.pdf}
\vspace{-0.1in}
\caption{\hl{Maps} %MDPI: Please change the hyphen (-) into a minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 generated by applying the trained chemical variable models to the hyperspectral data cubes collected on 23 November. Overlaid are color-filled squares showing the in situ reference data for the same collection period. The size of the squares is exaggerated for the visualization. We note that there is good agreement between the model predictions and reference data.\label{fig:map-chem}}
\end{figure}

>>>>

%=================================================================
% Full title of the paper (Capitalized)
 \Title{Characterizing Water Composition with an Autonomous Robotic Team Employing Comprehensive In Situ Sensing, Hyperspectral Imaging, Machine Learning, and Conformal Prediction}%AE title change

@@

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
Hanson Center for Space Sciences, University of Texas at Dallas, Richardson, TX 75080, USA;  john.waczak@utdallas.edu (J.W.); adam.aker@utdallas.edu (A.A.); lhw150030@utdallas.edu (L.O.H.W.); shawhintalebi@gmail.com (S.T.); ashen.fernando@utdallas.edu  (B.F.); pxh180012@utdallas.edu (P.M.H.D.); mazhar.iqbal@utdallas.edu (M.I.); MDL210001@utdallas.edu (M.L.); captdaveschaefer@gmail.com (D.S.)}

@@

\begin{figure}[H]
\vspace{-0.15in}
\includegraphics[width=0.9\columnwidth]{figures/results/fits/physical-fitres.png}
\vspace{-0.1in}
\caption{Scatter diagrams (\textbf{left}) and quantile--quantile plots (\textbf{right}) for the hyperparameter-optimized RFR models for the physical variables measured by the USV.\label{fig:physical-fit}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/fits/physical-ranking.pdf}
\vspace{-0.1in}
\caption{Ranked permutation importance for each feature in the physical variable models. Permutation importance measured the decrease in the model's $R^2$ value after replacing each feature in the prediction set with a random permutation of its values.\label{fig:physical-fi}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=0.9\columnwidth]{figures/results/maps/physical.pdf}
\vspace{-0.05in}
\caption{Maps generated by applying each of the physical variable models to the hyperspectral data cubes collected on 23 November. Overlaid over the predictions are color-filled squares showing the associated in situ reference data for the same collection period. The size of the squares has been exaggerated for visualization. We note that there is good agreement between the model predictions and the reference data. \label{fig:map-physical}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/fits/ions-fitres.png}
\vspace{-0.1in}
\caption{Scatter diagrams (\textbf{left}) and quantile--quantile plots (\textbf{right}) for the hyperparameter-optimized RFR models for the ion measurements made by the USV.\label{fig:ions-fit}}
\end{figure}

\vspace{-9pt}
\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\vspace{-0.1in}
\includegraphics[width=18 cm]{figures/results/fits/ions-ranking.pdf}
\vspace{-0.1in}
\end{adjustwidth}
\caption{Ranked permutation importance for the top 25 features of the ion models. The permutation importance measures the decrease in the model's $R^2$ value when each feature is replaced by a random permutation of its values.\label{fig:ions-fi}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/maps/ions.pdf}
\vspace{-0.1in}
\caption{Maps generated by applying the trained ion models to the data cubes collected on 23 November. Overlaid on the maps are the in situ reference measurements for the same collection period. The size of the squares has been exaggerated for the visualization. We note that there is good agreement between the generated map and the reference data. \label{fig:map-ions}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/fits/biochemical-fitres.png}
\vspace{-0.1in}
\caption{Scatter plots (\textbf{left}) and quantile--quantile plots (\textbf{right}) for the final hyperparameter-optimized models for the biochemical targets blue--green algae (phycoerythrin), CDOM, chlorophyll-a, and blue--green algae (phycocyanin). \label{fig:biochem-fit}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=0.9\columnwidth]{figures/results/fits/biochemical-ranking.pdf}
\vspace{-0.1in}
\caption{Ranked permutation importance for each feature in the trained biochemical models. The permutation importance measures the decrease in the model's $R^2$ value after replacing each feature with a random permutation of its values.\label{fig:biochem-fi}}
\end{figure}

@@

\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=0.92\columnwidth]{figures/results/maps/biochemical.pdf}
\vspace{-0.1in}
\caption{Maps generated 
 by applying the trained biochemical models to the data cubes collected on 23 November. Overlaid are the in situ reference data for the same collection period. The size of the squares has been exaggerated for the visualization. We note there is good agreement between the predicted map and the reference data. \label{fig:map-biochem}}
\end{figure}

@@

The final two models to consider are for the measured chemical concentrations of crude oil (CO) and optical brighteners (OB). The crude oil measurement includes natural unprocessed petroleum, whereas optical brighteners consist of whitening agents that are often added to products such as soaps, detergents, and cleaning agents. Both the crude oil and optical brightener measurements show multi-modal spatial distributions across each collection period. Scatter diagrams and quantile--quantile plots for the fitted models are shown in Figure~\ref{fig:chemicals-fit}. Both models achieve good performance, with $R^2$ values of 0.957 and 0.941 for CO and OB on the holdout test set. The performance of the CO model degrades for concentrations below 24 ppb, for which there are few records. Similarly, the OB model shows worse performance for concentrations below roughly 3.5 ppb.

@@

\vspace{-0.21in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/fits/chemical-fitres.png}
\vspace{-0.11in}
\caption{Scatter diagrams (\textbf{left}) and quantile--quantile plots (\textbf{right}) for the hyperparameter-optimized RFR models for the chemical variables measured by the USV.\label{fig:chemicals-fit}}
\end{figure}

\vspace{-9pt}
\begin{figure}[H]
\vspace{-0.15in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/fits/chemical-ranking.pdf}
\vspace{-0.15in}
\caption{Ranked permutation importance for the top 25 features of the chemical models. The permutation importance measures the decrease in the model's $R^2$ value after replacing each feature in the prediction set with a random permutation of its values.\label{fig:chemicals-fi}}
\end{figure}

@@

\vspace{-0.2in}
\hspace{-6pt}\includegraphics[width=\columnwidth]{figures/results/maps/chemical.pdf}
\vspace{-0.1in}
\caption{Maps generated by applying the trained chemical variable models to the hyperspectral data cubes collected on 23 November. Overlaid are color-filled squares showing the in situ reference data for the same collection period. The size of the squares is exaggerated for the visualization. We note that there is good agreement between the model predictions and reference data.\label{fig:map-chem}}
\end{figure}
